---
title: "Spotify Data"
author: Lucia Kajanová, Lukáš Poledňa, Tereza Slaninová, Kateřina Válková
date: 12. 12. 2023
output: ioslides_presentation
---
```{r setup, include=FALSE}
library(ggplot2)
library(MASS)
library(dplyr)
library(writexl)
library(tidyverse)
library(hrbrthemes)
library(ggpubr)
library(readxl)
library(cowplot)
library(vctrs)
library(psych)
library(tidyr)
```

# Introduction

## DATA
* Most Streamed Spotify Songs 2023
* Kaggle — Nidula Elgiriyewithana
* numeric variable  — streams




## DESCRIPTION OF VARIABLES

* PLAYLISTS – appearance of songs in different platforms’ playlists
* CHARTS – appearance of songs in different platforms’ charts
* MUSICAL ATTRIBUTES – beats per minute, key and mode of the song
* ARTIST INFORMATION – name of artist(s), number of artists contributing to songs
* SONG INFORMATION – track name, day, month and year of release
* OTHERS – danceability, acousticness, liveness, ...


## SUMMARY
```{r, include = FALSE}
spotify.2023_finalni_uprava <- read_excel("C:/Users/poled/OneDrive - MUNI/5. semestr/M5120/projekt/5120_projekt/spotify.2023_finalni_uprava.xlsx")
spotify_data <- spotify.2023_finalni_uprava
```
```{r}
summary(spotify_data[,1:8])
```
##
```{r}
summary(spotify_data[,9:16])
```

##
```{r}
summary(spotify_data[,17:24])
```

##
```{r, echo = FALSE}
p <- ggplot(spotify_data, aes(x = streams))
p <- p + geom_histogram(binwidth = 100000000, fill = "#1DB954", color = "#008222", alpha = 1000)
p <- p + xlab("Streams")
p <- p + ylab("Number of songs")
p <- p + theme(plot.title = element_text(size = 15))
p <- p + theme_bw()
p
```


##
```{r, echo = FALSE}
numericke <- spotify_data %>% select_if(is.numeric)
gr <- colorRampPalette(c("black", "white", "#1DB954"))
corPlot(numericke, gr = gr, main = "")
```

##
```{r, echo = FALSE}
gr2 <- colorRampPalette(c("black", "white", "#1DB954"))
corPlot(numericke[, 5:12], gr = gr2, main = "", upper = FALSE)
```

##
```{r, echo = FALSE}
pairs(spotify_data[, c(7, 8, 9, 10, 11, 12, 13, 14)], col="#1DB954", pch=19)
```

## SPOTIFY/APPLE/DEEZER PLAYLISTS
```{r, include = FALSE}
# Spotify
p11 <- ggplot(spotify_data, aes(x = in_spotify_playlists))
p11 <- p11 + geom_histogram(binwidth = 1000, fill = "#1DB954", color = "#008222", alpha = 1000)
p11 <- p11 + xlab("Spotify playlists")
p11 <- p11 + ylab("Number of songs")
p11 <- p11 + theme(plot.title = element_text(size = 15))
p11 <- p11 + theme_bw()
p11

# Apple
p33 <- ggplot(spotify_data, aes(x = in_apple_playlists))
p33 <- p33 + geom_histogram(binwidth = 20, fill = "#1DB954", color = "#008222", alpha = 1000)
p33 <- p33 + xlab("Apple Music playlists")
p33 <- p33 + ylab("Number of songs")
p33 <- p33 + theme(plot.title = element_text(size = 15))
p33 <- p33 + theme_bw()
p33

# Deezer
p22 <- ggplot(spotify_data, aes(x = in_deezer_playlists))
p22 <- p22 + geom_histogram(binwidth = 20, fill = "#1DB954", color = "#008222", alpha = 1000)
p22 <- p22 + xlab("Deezer playlists")
p22 <- p22 + ylab("Number of songs")
p22 <- p22 + theme(plot.title = element_text(size = 15))
p22 <- p22 + theme_bw()
p22
```

```{r, echo = FALSE}
p <- plot_grid(p11, p22, p33,
               ncol = 3, nrow = 1)
p
```

## SPOTIFY/APPLE/DEEZER CHARTS
```{r, include = FALSE, results = FALSE, warning = FALSE}
# Spotify
p4 <- ggplot(spotify_data, aes(x = in_spotify_charts))
p4 <- p4 + geom_histogram(binwidth = 10, fill = "#1DB954", color = "#008222", alpha = 1000)
p4 <- p4 + xlim(1, 150)
p4 <- p4 + xlab("Spotify charts - only ranked")
p4 <- p4 + ylab("Number of songs")
p4 <- p4 + theme(plot.title = element_text(size = 15))
p4 <- p4 + theme_bw()
p4

# Apple
p6 <- ggplot(spotify_data, aes(x = in_apple_charts))
p6 <- p6 + geom_histogram(binwidth = 10, fill = "#1DB954", color = "#008222", alpha = 1000)
p6 <- p6 + xlim(1, 290)
p6 <- p6 + xlab("Apple Music charts - only ranked")
p6 <- p6 + ylab("Number of songs")
p6 <- p6 + theme(plot.title = element_text(size = 15))
p6 <- p6 + theme_bw()
p6

# Deezer
p5 <- ggplot(spotify_data, aes(x = in_deezer_charts))
p5 <- p5 + geom_histogram(binwidth = 10, fill = "#1DB954", color = "#008222", alpha = 1000)
p5 <- p5 + xlim(1, 50)
p5 <- p5 + ylim(0, 100)
p5 <- p5 + xlab("Deezer charts - only ranked")
p5 <- p5 + ylab("Number of songs")
p5 <- p5 + theme(plot.title = element_text(size = 15))
p5 <- p5 + theme_bw()
p5
```
```{r, echo = FALSE, results = FALSE, warning = FALSE}
p <- plot_grid(p4, p6, p5,
               ncol = 3, nrow = 1)
p
```

## ARTISTS
```{r, echo = FALSE, results = FALSE}
levels(spotify_data$artists_name)

interpreti <- data.frame(table(spotify_data$artists_name))
interpreti_vyber <- subset(interpreti[interpreti$Freq >= 4,])

label_data <- interpreti_vyber
number_of_bar <- nrow(label_data)

numbers <- seq(from = 1, to = length(label_data$Var1))

angle <-  90 - 360 * (numbers-0.5) /number_of_bar

label_data$hjust<-ifelse( angle < -90, 1, 0)

label_data$angle<-ifelse(angle < -90, angle+180, angle)

p <- ggplot(interpreti_vyber, aes(x = as.factor(Var1), y = Freq))
p <- p + geom_bar(stat = "identity", fill = "#1DB954", alpha = 0.7)
p <- p + ylim(0, 34)
p <- p + theme_minimal()
p <- p + labs(x = NULL, y = NULL)
p <- p + theme(axis.ticks.x = element_blank())
p <- p + coord_polar(start = 0)
p
```








# Model selection

## MODELS
* Model 3A: $Streams_i = \beta_0 (Sch \times Ach)_i + \beta_1 (Ach \times Dch)_i \\ + \beta_2 (Sch \times Dch)_i + \beta_3 (Sch \times Dch \times Ach)_i$
* Model 3B: $Streams_i = \beta_0 + \beta_1 \times Sch + \beta_2 \times Ach + \beta_3 \times Dch \\+ \beta_4 \times Sp + \beta_5 \times Ap + \beta_6 \times Dp + \beta_7 \times (Sp \times Sch)_i \\+ \beta_8 \times (Ap \times Ach)_i + \beta_9 \times (Dch \times Dp)_i + \epsilon_i$

##
* Model 6: $Streams_i = \beta_0 + \beta_1 \times Sp + \beta_2 \times Ap + \beta_3 \times Dp \\+ \beta_4 \times (Sp \times Ap)_i + \beta_5 \times (Ap \times Dp)_i + \beta_6 \times (Sp \times Dp)_i \\+ \beta_7 \times (Sp \times Dp \times Ap)_i + \epsilon_i$
* Model 8: $Streams_i = \beta_0 + \beta_1 \times Sch + \beta_2 \times Ach + \beta_3 \times Dch \\+ \beta_4 \times Sp + \beta_5 \times Ap + \beta_6 \times Dp + \beta_7 \times (Sp \times Sch)_i + \\ \beta_8 \times (Ap \times Ach)_i + \beta_9 \times (Dch \times Dp)_i +\\ \beta_{10} \times (Sp \times Ap)_i + \beta_{11} \times (Ap \times Dp)_i \\+ \beta_{12} \times (Sp \times Dp)_i + \beta_{13} \times (Sp \times Dp \times Ap)_i + \epsilon_i$

##
* Model 9: $Streams_i = \beta_0 + \beta_1 \times Sch + \beta_2 \times Ach \\ + \beta_3 \times Sp + \beta_4 \times Ap + \beta_5 \times Dp  \\ + \beta_6 \times (Sp \times Dp)_i + \beta_7 \times (Sp \times Ap)_i + \\ \beta_8 \times (Dp \times Ap)_i +
 \beta_9 \times (Ap \times Ach)_i + \beta_{10} \times (Sp \times Sch)_i + \\ \beta_{11} \times (Sp \times Dp \times Ap)_i+ \epsilon_i$
* Model 11: $Streams_i = \beta_0 + \beta_1 \times Sch   + \beta_2 \times Sp + \beta_3 \times Ap + \\ \beta_4 \times Dp + \beta_5 \times (Sp \times Dp)_i + \\ \beta_6 \times (Sp \times Ap)_i +  \beta_7 \times (Dp \times Ap)_i \\+
\beta_8 \times (Sp \times Sch)_i +  \beta_9 \times (Sp \times Dp \times Ap)_i + \epsilon_i$

##
```{r, include = FALSE, results = FALSE}
# Model 3A
model3a <- lm(streams ~ in_spotify_charts * in_apple_charts + in_apple_charts* in_deezer_charts + in_spotify_charts *in_deezer_charts + in_spotify_charts* in_deezer_charts * in_apple_charts, data = spotify_data )
summary(model3a) # 0.1325
AIC(model3a) #34954.95
BIC(model3a) #34997.43

# Model 3B
model3b <- lm(streams ~ in_spotify_playlists*in_spotify_charts + in_apple_playlists* in_apple_charts + in_deezer_charts *in_deezer_playlists, data = spotify_data)
summary(model3b) #0.6513
AIC(model3b) # 34201.31
BIC(model3b) # 34253.23

# Model 6
model6 <- lm(streams ~ in_spotify_playlists*in_apple_playlists + in_apple_playlists* in_deezer_playlists + in_spotify_playlists *in_deezer_playlists + in_spotify_playlists* in_deezer_playlists*in_apple_playlists, data = spotify_data )
summary(model6) # 0.6776
AIC(model6) # 34134.38
BIC(model6) # 34176.86

# Model 8
model8 <- lm(streams ~ in_spotify_playlists*in_spotify_charts + in_apple_playlists* in_apple_charts + in_deezer_charts *in_deezer_playlists + in_spotify_playlists*in_apple_playlists + in_apple_playlists* in_deezer_playlists + in_spotify_playlists *in_deezer_playlists + in_spotify_playlists* in_deezer_playlists*in_apple_playlists, data = spotify_data)
summary(model8) # 0.6921
anova(model8)
AIC(model8) # 34102.16
BIC(model8) # 34172.96

# Model 9
model9 <- lm(streams ~ in_spotify_playlists* in_deezer_playlists*in_apple_playlists + in_spotify_playlists *in_deezer_playlists+ in_apple_playlists* in_apple_charts + in_spotify_playlists*in_spotify_charts + in_apple_playlists +in_spotify_playlists+in_spotify_charts,  data = spotify_data)
summary(model9)# 0.6922
anova(model9)
AIC(model9) # 34100.04
BIC(model9) # 34161.4

# Model 11
model11 <- lm(streams ~ in_spotify_playlists* in_deezer_playlists*in_apple_playlists + in_spotify_playlists *in_deezer_playlists+ in_spotify_playlists*in_spotify_charts + in_apple_playlists +in_spotify_playlists+in_spotify_charts,  data = spotify_data)
summary(model11) # 0.6922
anova(model11)
AIC(model11) # 34097.86
BIC(model11) # 34149.11

data <- data.frame(matrix(
  c(0.1325, 0.6513, 0.6776, 0.6921, 0.6922, 0.6922, AIC(model3a), AIC(model3b), AIC(model6), AIC(model8),
             AIC(model9), AIC(model11), BIC(model3a), BIC(model3b), BIC(model6), BIC(model8), BIC(model9), BIC(model11)),
  ncol = 6,
  nrow = 3,
  byrow = TRUE
)
)
rownames(data) <- c("Adjusted R squared", "AIC", "BIC")
colnames(data) <- c("Model 3A", "Model 3B", "Model 6", "Model 8", "Model 9", "Model 11")
```
```{r, echo = FALSE}
print(data)
```

## MODEL 9
```{r, echo = FALSE}
#Model 9
plot(studres(model9)~c(1:829),
     ylab="Jackknife residuals", xlab="Index", main="Jackknife residuals")
abline(h=0, col="#1DB954", lwd=2)
abline(h = qt(1-0.05/(2 * 829), df=817), col="#008222")
abline(h = -qt(1-0.05/(2 * 829), df=817), col="#008222")
indices <- c(527, 528, 42, 37, 136, 721)
abline(v=indices, lty=2)
points(indices, studres(model9)[indices], pch = 16, col = 'red')

```


## MODEL 11
```{r, echo = FALSE}
# Model 11
plot(studres(model11)~c(1:829),
     ylab="Jackknife residuals", xlab="Index", main="Jackknife residuals")
abline(h=0, col="#1DB954", lwd=2)
abline(h = qt(1-0.05/(2 * 829), df=819), col="#008222")
abline(h = -qt(1-0.05/(2 * 829), df=819), col="#008222")

indices <- c(527, 528, 42, 37, 136, 721)

abline(v=indices, lty=2)
points(indices, studres(model11)[indices], pch = 16, col = 'red')
```

##
```{r, include = FALSE, results = FALSE}
# Model 9
spotify_data_bez_hodnot_model9 <- spotify_data[c(-527, -42, -528, -721),]
model9_1 <- lm(streams ~ in_spotify_playlists* in_deezer_playlists*in_apple_playlists + in_spotify_playlists *in_deezer_playlists+ in_apple_playlists* in_apple_charts + in_spotify_playlists*in_spotify_charts + in_apple_playlists +in_spotify_playlists+in_spotify_charts,  data = spotify_data_bez_hodnot_model9)
summary(model9_1) # 0.7233
anova(model9_1)
AIC(model9_1) # 33806.12
BIC(model9_1) # 33867.42


# Model 11
spotify_data_bez_hodnot_model11 <- spotify_data[c(-527, -42, -528, -721),]
model11_1 <- lm(streams ~ in_spotify_playlists* in_deezer_playlists*in_apple_playlists + in_spotify_playlists *in_deezer_playlists+ in_spotify_playlists*in_spotify_charts + in_apple_playlists +in_spotify_playlists+in_spotify_charts,  data = spotify_data_bez_hodnot_model11)
summary(model11_1) # 0.7221
anova(model11_1)
AIC(model11_1) # 33807.87
BIC(model11_1) # 33859.74

data_2 <- data.frame(matrix(
  c(0.7233, 0.7221, AIC(model9_1), AIC(model11_1), BIC(model9_1), BIC(model11_1)),
  ncol = 2,
  nrow = 3,
  byrow = TRUE
)
)
rownames(data_2) <- c("Adjusted R squared", "AIC", "BIC")
colnames(data_2) <- c("Model 9_1", "Model 11_1")
```
```{r, echo = FALSE}
print(data_2)
```


# Model 9

##
```{r, echo = FALSE}
par(mfrow = c(2, 3))
plot(model9, which = 1:6)
```

##
```{r, echo = FALSE}
plot(studres(model9)~c(1:829),
     ylab="Jackknife residuals", xlab="Index", main="Jackknife residuals")
abline(h=0, col="#1DB954", lwd=2)
abline(h = qt(1-0.05/(2 * 829), df=817), col="#008222")
abline(h = -qt(1-0.05/(2 * 829), df=817), col="#008222")

indices <- c(527, 528, 42, 37, 136, 721)

abline(v=indices, lty=2)
points(indices, studres(model9)[indices], pch = 16, col = 'red')
```

## MODEL 9_1
```{r, include = FALSE}
spotify_data_bez_hodnot_model9 <- spotify_data[c(-527, -42, -528, -721),]
model9_1 <- lm(streams ~ in_spotify_playlists* in_deezer_playlists*in_apple_playlists + in_spotify_playlists *in_deezer_playlists+ in_apple_playlists* in_apple_charts + in_spotify_playlists*in_spotify_charts + in_apple_playlists +in_spotify_playlists+in_spotify_charts,  data = spotify_data_bez_hodnot_model9)
```
```{r, echo = FALSE}
plot(studres(model9_1)~c(1:825),
     ylab="Jackknife residuals", xlab="Index", main="Jackknife residuals")
abline(h=0, col="#1DB954", lwd=2)
abline(h = qt(1-0.05/(2 * 825), df=813), col="#008222")
abline(h = -qt(1-0.05/(2 * 825), df=813), col="#008222")
indices <- c(125, 107, 145, 38, 37)

abline(v=indices, lty=2)
points(indices, studres(model9_1)[indices], pch = 16, col = "red")
```

## MODELS 9, 9_1, 9_2
```{r, include = FALSE}
spotify_data_bez_hodnot2_model9 <- spotify_data_bez_hodnot_model9[c(-125, -145),]
model9_2 <- lm(streams ~ in_spotify_playlists* in_deezer_playlists*in_apple_playlists + in_spotify_playlists *in_deezer_playlists+ in_apple_playlists* in_apple_charts + in_spotify_playlists*in_spotify_charts + in_apple_playlists +in_spotify_playlists+in_spotify_charts,  data = spotify_data_bez_hodnot2_model9)
```
```{r, echo = FALSE}
data_3 <- data.frame(matrix(
  c(0.6922, 0.7233, 0.7324, AIC(model9), AIC(model9_1), AIC(model9_2), BIC(model9), BIC(model9_1), BIC(model9_2), 0.2956, 0.001073, 0.0004947),
  ncol = 3,
  nrow = 4,
  byrow = TRUE
)
)
rownames(data_3) <- c("Adjusted R squared", "AIC", "BIC", "Shapiro Test")
colnames(data_3) <- c("Model 9", "Model 9_1", "Model 9_2")
print(data_3)
```

##
```{r, echo = FALSE}
par(mfrow = c(2, 3))
plot(model9_2, which = 1:6)
```

# Prediction intervals

## MODEL 1
```{r, echo = FALSE}
model1 <- lm(formula = streams ~ in_spotify_playlists, data = spotify_data)

plot(spotify_data$streams ~ spotify_data$in_spotify_playlists, xlab = "Spotify Playlists", ylab = "Streams")
xx <- seq(min(spotify_data$in_spotify_playlists), max(spotify_data$in_spotify_playlists), length=501)
novadata1 <- data.frame(in_spotify_playlists = xx)
yy1 <- predict(model1, newdata = novadata1, interval = "prediction")

lines(xx, yy1[, 1], col="#1DB954", lwd=2)
lines(xx, yy1[, 2], col="#008222", lwd=2, lty=3)
lines(xx, yy1[, 3], col="#008222", lwd=2, lty=3)
```

## MODEL 2
```{r, echo = FALSE}
model2 <- lm(formula = streams ~ in_apple_playlists, data = spotify_data)

plot(spotify_data$streams ~ spotify_data$in_apple_playlists, xlab = "Apple Playlists", ylab = "Streams")
xx2 <- seq(min(spotify_data$in_apple_playlists), max(spotify_data$in_apple_playlists), length=501)
novadata2 <- data.frame(in_apple_playlists = xx2)

yy2 <- predict(model2, newdata = novadata2, interval = "prediction")

lines(xx2, yy2[, 1], col="#1DB954", lwd=2)
lines(xx2, yy2[, 2], col="#008222", lwd=2, lty=3)
lines(xx2, yy2[, 3], col="#008222", lwd=2, lty=3)
```

# Confidence intervals

## MODEL 1
```{r, echo = FALSE}
model1 <- lm(formula = streams ~ in_spotify_playlists, data = spotify_data)

plot(spotify_data$streams ~ spotify_data$in_spotify_playlists, xlab = "Spotify Playlists", ylab = "Streams")
xx <- seq(min(spotify_data$in_spotify_playlists), max(spotify_data$in_spotify_playlists), length=501)
novadata1 <- data.frame(in_spotify_playlists = xx)
yy <- predict(model1, newdata = novadata1, interval = "confidence")

lines(xx, yy[, 1], col="#1DB954", lwd=2)

lines(xx,
      yy[, 1] - (yy[, 3] - yy[, 1])*sqrt(2*qf(0.95, df1=2, df2=827))/qt(0.975, df=827),
      col="#008222", lwd=2, lty=2)
lines(xx,
      yy[, 1] + (yy[, 3] - yy[, 1])*sqrt(2*qf(0.95, df1=2, df2=827))/qt(0.975, df=827),
      col="#008222", lwd=2, lty=2)
var.bands <- diag(2*qf(0.95, df1=2, df2=827)*cbind(1, xx)%*%vcov(model1)%*%rbind(1, xx))
lines(xx, yy[, 1]+sqrt(var.bands), col="#008222", lwd=2, lty=2)
lines(xx, yy[, 1]-sqrt(var.bands), col="#008222", lwd=2, lty=2)
```

## MODEL 2
```{r, echo = FALSE}
model2 <- lm(formula = streams ~ in_apple_playlists, data = spotify_data)

plot(spotify_data$streams ~ spotify_data$in_apple_playlists, xlab = "Apple Playlists", ylab = "Streams")
xx2 <- seq(min(spotify_data$in_apple_playlists), max(spotify_data$in_apple_playlists), length=501)
novadata2 <- data.frame(in_apple_playlists = xx2)
yy2 <- predict(model2, newdata = novadata2, interval = "confidence")

lines(xx2, yy2[, 1], col="#1DB954", lwd=2)

lines(xx2,
      yy2[, 1] - (yy2[, 3] - yy2[, 1])*sqrt(2*qf(0.95, df1=2, df2=827))/qt(0.975, df=827),
      col="#008222", lwd=2, lty=2)
lines(xx2,
      yy2[, 1] + (yy2[, 3] - yy2[, 1])*sqrt(2*qf(0.95, df1=2, df2=827))/qt(0.975, df=827),
      col="#008222", lwd=2, lty=2)
var.bands <- diag(2*qf(0.95, df1=2, df2=827)*cbind(1, xx2)%*%vcov(model2)%*%rbind(1, xx2))
lines(xx2, yy2[, 1]+sqrt(var.bands), col="#008222", lwd=2, lty=2)
lines(xx2, yy2[, 1]-sqrt(var.bands), col="#008222", lwd=2, lty=2)
```


# Thank you for your \ attention
